### 2024-03-07 paper_version/rebuttal_letter.tex
review

%-----------------------------------------------------------------------------%
\documentclass[a4paper, 10pt]{letter}

\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{xcolor}
\geometry{a4paper,scale=0.8}

%-----------------------------------------------------------------------------%

\begin{document}

\begin{letter}
\opening{Dear Editors,}

We thank the reviews for the generous and insightful comments on our work and have revised the manuscript to address their concerns. The corresponding replies for each comments are attached in the end.

We believe that the manuscript is now suitable for publication.

\closing{On behalf of all authors,\\ \vspace{1cm} Tianwei Yu}
\end{letter}

\newpage

\begin{itemize}
  \item Question 1.1: \textcolor{violet}{Can the new 3D scheme reduces to Degond's 1D scheme?}
  \item[$\blacktriangleright$] Reply: \textcolor{blue}{
  Yes. The proposed numerical scheme in 3D is a natural extension of the 1D
  the scheme proposed by Degond et. al. This is justified by considering problems with
  transverse invariance (namely, variables keep constant along the x-axis and y-axis) and
  verifying that the "projected" 3D scheme coincides with the original 1D scheme at the z-axis.
  Section~4.1 is devoted to such a purpose. To better clarify this point, we add a remark
  right after the 3D full discretization: ``When applying the ``projection to one dimension"
  described in Section~4.1.1, the fully discrete 3D AP scheme (32) collapses to (31)
  discretization proposed in $[$P.~Degond, F.~Deluzet, and D.~Savelief, \emph{Numerical approximation of the Euler-Maxwell model in the quasineutral limit}, Journal of Computational Physics, 231 (4), pp.~1917--1946, 2012$]$. This is not a coincidence, because
  recovering of (31) under dimensional reduction has guided our construction."
  }

  \item Question 1.2: \textcolor{violet}{In Fig 8, Page 19, it seems the 1D results are better than the 3D results by using the same meshes. Can the new 3D scheme
  preserve some good properties of the original 1D scheme?}
  \item[$\blacktriangleright$] Reply: \textcolor{blue}{
  Technically speaking, Fig.~8 is a ``verification" rather than a ``comparison".
  The result produced by the 3D scheme, as illustrated in Section~5.1, is the
  solution along the z-axis for the cylindrical setting in Fig.~2. To better
  clarify this point, we add some comments in the description of Fig.~8:
  ``... The 3D result is generated under the setting in Fig.~2
  and restricted to the z-axis. The results are slightly different from
  the 1D result probably due to the impact of the outer boundary of the
  cylindrical domain."
  }

  \item Question 1.3: \textcolor{violet}{In numerical tests, the mass ratio and the collision parameter are computed by a constant number. Is it sufficient to
demonstrate the good performance of the new scheme?}
  \item[$\blacktriangleright$] Reply: \textcolor{blue}{
  We agree that the robustness with respect to the mass ratio and the collision parameter
  is worth studying in dealing with the plasma model.
  However, the objective of the proposed scheme is stability with respect to
  the Debye length $\lambda$. The impacts of the mass ratio and the collision parameter
  are not of our central interest.
  In fact, our scheme does not possess uniform stability over the mass ratio $\varepsilon_*$
  yet has it over the collision parameter $\alpha^{\text{coll}}$. This can be
  inferred from (32j) by noticing the explicit term
  $\varepsilon_*^{-2}q_*(n\mathbf{u})_*^m \times \mathbf{B}^m$
  and the implicit terms
  $\varepsilon_*^{-2}\mathbf{R}^{\text{coll}, m+1}_*, \varepsilon_*^{-2}Q^{\text{coll}, m+1}_*$.
  It seems possible to achieve stability over the mass ratio $\varepsilon_*$
  by replacing $\varepsilon_*^{-2}q_*(n\mathbf{u})_*^m \times \mathbf{B}^m$
  by $\varepsilon_*^{-2}q_*(n\mathbf{u})_*^{m+1} \times \mathbf{B}^{m}$.
  However, this would significantly change the time-stepping procedure.
  We decide not to pursue this and want to be aligned with Degond's scheme.
  }

  \item Question 1.4: \textcolor{violet}{Include the following references:
  [Fast-Converging and Asymptotic-Preserving Simulation of Frequency Domain Thermoreflectance, Jia Liu \& Lei Wu. Commun. Comput. Phys., 34 (2023), pp. 65-93.]
  [A Coupled FEM-BEM Approach for the Solution of the Free-Boundary Axi-Symmetric Plasma Equilibrium Problem. M. Bonotto, D. Abate, P. Bettni \& F. Villone. Commun. Comput. Phys., 31 (2022), pp. 27-59.]}
  \item[$\blacktriangleright$] Reply: \textcolor{blue}{Both paper is included in the introduction.
  The first paper is added to the literature review on AP schemes
  while the second one is mentioned when the symmetry of the problem is discussed.}
\end{itemize}

\newpage

\begin{itemize}
  \item Question 2.1: \textcolor{violet}{The literature review provided in the paper is incomplete.
  Numerous structure-preserving schemes for plasma models involving the Euler
  and Maxwell equations exist, such as the compressible ideal MHD model.
  Relevant works should include, but not be limited to:
  \begin{itemize}
      \item[1)] The positivity-preserving and divergence-free finite volume schemes and
  central discontinuous Galerkin schemes (also using the primal and dual meshes);
  see, e.g., [K. Wu, SIAM Journal on Numerical Analysis, 56(4):2124-2147, 2018]
  and [K. Wu, H. Jiang, and C.-W. Shu, SIAM Journal on Numerical Analysis, 61: 250-285, 2023].
      \item[2)] These schemes are designed by using the geometric quasilinearization framework in
  [K. Wu and C.-W. Shu, SIAM Review, 65(4):1031-1073, 2023],
  where the multicomponent MHD equations for multi-species plasma were also studied.
      \item[3)] High-order asymptotic preserving (AP) finite difference WENO schemes with
  constrained transport were also developed for MHD equations in
  [W. Chen, K. Wu, and T. Xiong, Journal of Computational Physics, 488: 112240, 2023].
  \end{itemize}}

  \item[$\blacktriangleright$] Reply: \textcolor{blue}{
  We have added a paragraph at the end of Section~2.2. A short discussion and literature
  review on the numerical methods for plasma models is presented with a focus on the
  structure-preserving property. The literature mentioned in 3) is included in the
  literature review in Section~1 when discussing AP schemes.
  }

  \item Questions 2.2: \textcolor{violet}{The author did not provide an accuracy test
  for the convergence rate of their proposed numerical method.
  An illustrative example can be found in section 6.2 of
  [K. Wu, H. Jiang, and C.-W. Shu, SIAM Journal on Numerical Analysis, 61: 250-285, 2023],
  as well as in Examples 5.1 and 5.2 in [W. Chen, K. Wu, and T. Xiong, Journal of Computational Physics, 488: 112240, 2023]}
  \item[$\blacktriangleright$] Reply: \textcolor{blue}{
  To handle this issue, we put some effort into re-devising the matrix assembly which
  now admits a fast iterative solver such that reference solutions on sufficiently refined
  meshes are possible. A new case involving a cuboid domain is tested and
  $L^1$ and $L^2$ errors at different values of $\lambda$ are reported in the same manner as suggested by the second paper.
  A smooth test case seems hard to construct due to the presence of the interface of two different
  electromagnetic boundary conditions in our underlying setting depicted in Fig.~2.
  }

  \item Question 2.3: \textcolor{violet}{
  Limited Numerical Examples: All the numerical examples provided in the paper
  are either one-dimensional or three-dimensional cylindrical. To strengthen the paper's contribution,
  it is advisable to include a genuinely three-dimensional example.
  This is especially relevant as one of the primary contributions of the paper lies in the
  extension of a one-dimensional scheme to three dimensions.
  }
  \item[$\blacktriangleright$] Reply: \textcolor{blue}{
  We include a slightly more general 3D case in Case IV. Yet we would like to mention that 
  our current framework of primal-dual mesh generation is based on the idea of
  ``extruding 2D mesh" in Sec.~4.2.1 of
  $[$Marrone, Massimiliano.\emph{Computational Aspects of the Cell Method in Electrodynamics - Abstract.} Journal of Electromagnetic Waves and Applications 15 (2001): 407 - 408.$]$.
  It has not been extended yet to arbitrary geometries, which possibly takes quite some effort.
  Therefore, we have decided to defer this part to future work. Nevertheless,
  the proposed scheme can be directly applied to arbitrary domains
  as long as the primal-dual meshes are generated properly.
  }
\end{itemize}

\end{document}
%-----------------------------------------------------------------------------%